---
title: "Precarious employment - descriptive analysis"
author: "Andrew Pulford"
date: ""
output: word_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## remove any existing objects from global environment
rm(list=ls()) 

## stop scientific notation
options(scipen=999, digits = 1)

### load packages
library(tidyverse)
library(DiagrammeR)
library(knitr)
library(TraMineR)
remotes::install_github("rstudio/webshot2")

##### load dataframes ----------------------------------------------------------

### master raw data --------
# This includes selected variables for all individual respondents across Understanding Society waves 3-10
master_raw <- readRDS("../raw_data/master_raw1.rds")

## waves 3-6
master_raw_a <- master_raw %>% filter(wv_n %in% c(3:6))

## waves 7-10
master_raw_b <- master_raw %>% filter(wv_n %in% c(7:10))

### non-working age ---------
non_working_age_spine_a <- readRDS("../look_ups/non_working_age_spine_a.rds")
non_working_age_spine_b <- readRDS("../look_ups/non_working_age_spine_b.rds")

### no valid case at end-point ----------
no_endpoint_a <- readRDS("../look_ups/no_endpoint_a.rds")
no_endpoint_b <- readRDS("../look_ups/no_endpoint_b.rds")

### eligible population ---------
# This includes individual respondents aged 16-64; Understanding Society waves 3-10 
eligible_a <- readRDS("../working_data/eligible_pop_a.rds") 
eligible_b <- readRDS("../working_data/eligible_pop_b.rds") 

### censored cases ---------
censor_combined_a <- readRDS("../look_ups/censor_combined_a.rds")
censor_combined_b <- readRDS("../look_ups/censor_combined_b.rds")

### analytic samples  ---------
dfas1a <- read_rds("../analytic_sample_data/dfas1a.rds")
dfas1b <- read_rds("../analytic_sample_data/dfas1b.rds")


##### flowchart nodes A to xx --------------------------------------------------

### Node A:  ----------------
### Total number of individuals and observations in study period
## number of unique individuals (node A)
raw_unique_indiv_a <- length(unique(master_raw_a$pidp))
raw_unique_indiv_b <- length(unique(master_raw_b$pidp))

## number of observations (node A)
raw_observations_a <- nrow(master_raw_a)
raw_observations_b <- nrow(master_raw_b)
#raw_observations <- prettyNum(raw_observations, big.mark=",")

### Node AA: ----------------
### number of individuals excluded due to not being working-age or no valid 
### response at study endpoint

## non-working age
non_work_age_a <- nrow(non_working_age_spine_a)
non_work_age_b <- nrow(non_working_age_spine_b)
#non_work_age <- prettyNum(non_work_age, big.mark=",")

## no valid endpoint response
no_valid_end_a <- length(unique(no_endpoint_a$pidp))
no_valid_end_b <- length(unique(no_endpoint_b$pidp))
#no_wv2 <- prettyNum(no_wv2, big.mark=",")


### Node B: ----------------
### number of individuals and observations in eligible population
eligible_unique_indiv_a <- length(unique(eligible_a$pidp))
eligible_unique_indiv_b <- length(unique(eligible_b$pidp))
#eligible_descr_unique_indiv <- prettyNum(eligible_descr_unique_indiv, big.mark=",")


### Node BB: ----------------
### number of individuals censored due to retirement or incomplete

## retired
retired_spine_a <- censor_combined_a %>% 
  filter(censor_reason=="retired")

retired_spine_b <- censor_combined_b %>% 
  filter(censor_reason=="retired")

censor_retired_a <- nrow(retired_spine_a)
censor_retired_b <- nrow(retired_spine_b)
#censor_retired <- prettyNum(censor_retired, big.mark=",")


## number of individuals censored due to incomplete data
incomplete_spine_a <- censor_combined_a %>% 
  filter(censor_reason=="incomplete")

incomplete_spine_b <- censor_combined_b %>% 
  filter(censor_reason=="incomplete")

incomplete_a <- nrow(incomplete_spine_a)
incomplete_b <- nrow(incomplete_spine_b)
#censor_lost <- prettyNum(censor_lost, big.mark=",")


### Node C: ---------------- 
### Number of individuals and observations in analytic samples

dfas1a_observations <- nrow(dfas1a)
dfas1b_observations <- nrow(dfas1b)
#dfas1_observations <- prettyNum(dfas1_observations, big.mark=",")

dfas1a_unique_indivs <- length(unique(dfas1a$pidp))
dfas1b_unique_indivs <- length(unique(dfas1b$pidp))
#dfas1_unique_indivs <- prettyNum(dfas1_unique_indivs, big.mark=",")


##### Study endpoint sample characteristics
sample_chars_endpoint <- readRDS("../working_data/sample_chars_endpoint.rds")  %>% 
  mutate(measure=tolower(measure))

## formatting for table
# for estimates set 1dp
sample_chars_endpoint$est <- formatC(sample_chars_endpoint$est, digits = 1, format = "f")

# for number of cases set 1,000 separator
sample_chars_endpoint$n <- prettyNum(sample_chars_endpoint$n, big.mark=",")

# for number of cases set <10 to -999 so can suppress low numbers
sample_chars_endpoint <- sample_chars_endpoint %>% 
  mutate(n = ifelse(n%in%c(1:10),-999,n))




sample_chars_endpoint_wide <- sample_chars_endpoint %>% 
  filter(wv_n==6) %>% 
  rename("n6" = "n",
         "est6" = "est") %>% 
  ungroup() %>% 
  select(-wv_n)

temp <- sample_chars_endpoint %>% 
  filter(wv_n==10) %>% 
  rename("n10" = "n",
         "est10" = "est") %>% 
  ungroup() %>% 
  select(-wv_n)

sample_chars_endpoint_wide <- sample_chars_endpoint_wide %>% 
  full_join(temp, by = c("var", "measure"))

# for number of cases convert to character so NAs can be removed
sample_chars_endpoint_wide$n6 <- as.character(sample_chars_endpoint_wide$n6)
sample_chars_endpoint_wide$n10 <- as.character(sample_chars_endpoint_wide$n10)

sample_chars_endpoint_wide <- sample_chars_endpoint_wide %>% 
  group_by(var) %>% 
  mutate(row_n = row_number()) %>% 
  mutate(var = ifelse(row_n==1,var,"")) %>% 
  ungroup() %>% 
  mutate(n6 = ifelse(n6=="NA","",
                    ifelse(n6=="-999","-",n6)),
         n10 = ifelse(n10=="NA","",
                    ifelse(n10=="-999","-",n10))) %>% 
  select(-row_n)

#### Node E: -------------
### Outcome caseness

srh_case <- "x"
srh_nocase <- "x" 
ghq_case <- "x"
ghq_nocase <- "x"

```

## Sample description

A total of `r prettyNum(raw_unique_indiv, big.mark=",")` individuals participated in UKHLS waves 2-10 (Figure 1). We excluded `r prettyNum(non_work_age, big.mark=",")` individuals for not being working-age (16-64) for the full study period and `r prettyNum(no_wv2, big.mark=",")` individuals for not having a valid survey response at wave 2. From an eligible population of `r prettyNum(eligible_descr_unique_indiv, big.mark=",")` we excluded `r prettyNum(censor_retired, big.mark=",")` individuals who were retired during the study period, `r prettyNum(censor_death, big.mark=",")` who died and `r prettyNum(censor_lost, big.mark=",")` who did not have a valid response at wave 10. Table 1 presents key characteristics of the analytic sample at baseline by employment contract status (UKHLS wave 2).


**Figure 1: Flowchart for inclusion and exclusion of UKHLS respondents (waves 2-10) to create analytic sample**
```{r, echo=FALSE, include=TRUE, fig.width=8}


## flowchart - need to fix box AA to include no valid wave 2 response
grViz(paste0("digraph {
  graph []
  node []
    A [label = '",prettyNum(raw_unique_indiv_a, big.mark=",")," individuals
    
    (",prettyNum(raw_observations_a, big.mark=",")," observations)', shape = box, width = 3, height = 1]
    AA [label = '",prettyNum(non_work_age_a, big.mark=",")," non-working age
    
    ",prettyNum(no_valid_end_a, big.mark=",")," no valid wave 6 response' , shape = box, 
    width = 3, height = 1]
    B [label = '",prettyNum(eligible_unique_indiv_a, big.mark=",")," eligible individuals' , shape = box, 
    width = 3, height = 1]
    BB [label = '",prettyNum(censor_retired_a, big.mark=",")," retired
    
    ",prettyNum(incomplete_a, big.mark=",")," incomplete data', shape = box, width = 3, height = 1]
    C [label = '",prettyNum(dfas1a_unique_indivs, big.mark=",")," in analytic sample A
    
    (",prettyNum(dfas1a_observations, big.mark=",")," observations)', shape = box, 
    width = 3, height = 1]
    D [label = 'x in each employment history cluster' , shape = box, 
    width = 3, height = 1]
    E [label = '",prettyNum(srh_case, big.mark=",")," cases - poor self-reported health
    
    ",prettyNum(srh_nocase, big.mark=",")," event free - poor self-reported health
    
    ",prettyNum(ghq_case, big.mark=",")," cases - common mental health disorder
    
    ",prettyNum(ghq_nocase, big.mark=",")," event free - common mental health disorder' , shape = box, width = 3, height = 1]
  edge []
    A->B
    A->AA
    B->BB
    B->C
    C->D
    D->E
{ rank = same; A; AA }
{ rank = same; B; BB }}"))
```

```{r, echo=FALSE}
## add checks to ensure number of individuals adds up
#doesn't current
#check_AB <- eligible_descr_unique_indiv+non_work_age+no_wv2==raw_unique_indiv
#
#check_BC <- censor_death+censor_retired+censor_lost+dfas1_unique_indivs==eligible_descr_unique_indiv
#
#if(check_AB!=TRUE){print("boxes AA and B to not equal box A")}
#if(check_BC!=TRUE){print("boxes BB and C to not equal box A")}

```
\newpage
**Table 1: Analytic sample characteristics at baseline (UKHLS wave 2)**
```{r, echo=FALSE}

#sample_chars_endpoint_wide <- sample_chars_endpoint_wide[,c(1:10)]
kable(sample_chars_endpoint_wide, 
      col.names = c("Variable", 
                    "Measure", 
                    "n (wave 6)", 
                    "% (wave 6)", 
                    "n (wave 10)", 
                    "% (wave 10)"), 
      format.args = list(big.mark = ","))

```
\newpage
## Missing data
```{r, include=FALSE, echo=FALSE}
### total number of cases in each wave 16-64
spine_total <- dfas1 %>% group_by(wv_n) %>% summarise(n=n()) %>% ungroup()

## total number of cases by number of waves reported
spine_waves <- dfas1 %>% 
  group_by(pidp) %>% 
  summarise(n_wvs = n()) %>% # calculate number of waves responded to
  ungroup() %>% 
  group_by(n_wvs) %>% 
  summarise(n = n()) %>% # calculate number of cases
  ungroup() %>% 
  mutate(pc= n/sum(n)*100) # calculate percentage of cases

### sequence analysis
## wide format for sequencing data
spine_wide <- dfas1 %>%
  select(pidp,wv_n) %>% 
  mutate(response=1) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = response, values_fill = 99)

spine_labs <- c("response", "missing")
spine_code <- c("RESP", "NA")
spine.seq <- seqdef(spine_wide, 2:10, states = spine_code,
                    labels = spine_labs)


```
Responses to all nine waves in the study period were provided by `r spine_waves$pc[spine_waves$n_wvs==9]`% of the analytic sample (Figure 2). Response sequences ascending in order of frequency are presented in Figure 3.

**Figure 2: Number of waves responded to by percentage of the analytic sample**
```{r, echo=FALSE, fig.width=3,fig.height=3}
# chart
spine_waves %>% 
  ggplot(aes(x=n_wvs, y=pc)) +
  geom_col() +
  scale_x_continuous(limits = c(1,11),breaks = 2:9) +
  xlab("Number of waves responded to") +
  ylab("Percentage of cases") +
  theme_classic()
```

\newpage
**Figure 3: UKHLS response sequences ascending in order of frequency (waves 2-10)**

```{r, echo=FALSE, fig.width=12,fig.height=15}

# sequence frequency plot (all common sequences)
seqfplot(spine.seq, 
         idxs=1:900, # to add more lines
         with.legend = "right", 
         border = NA, 
         main = "Sequence frequency plot")
```

```{r, echo=FALSE, include=FALSE}


```


```{r, echo=FALSE, fig.width=9,fig.height=9}

#


```