---
title: "Precarious employment - descriptive analysis"
author: "Andrew Pulford"
date: ""
output: word_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## remove any existing objects from global environment
rm(list=ls()) 

## stop scientific notation
options(scipen=999, digits = 1)

### load packages
library(tidyverse)
library(DiagrammeR)
library(knitr)
library(TraMineR)
remotes::install_github("rstudio/webshot2")

##### load dataframes ----------------------------------------------------------

### master raw data --------
# This includes selected variables for all individual respondents across Understanding Society waves 3-10
master_raw <- readRDS("../raw_data/master_raw1.rds")

## waves 3-6
master_raw_a <- master_raw %>% filter(wv_n %in% c(3:6))

## waves 7-10
master_raw_b <- master_raw %>% filter(wv_n %in% c(7:10))

### non-working age ---------
non_working_age_spine_a <- readRDS("../look_ups/non_working_age_spine_a.rds")
non_working_age_spine_b <- readRDS("../look_ups/non_working_age_spine_b.rds")

### no valid case at end-point ----------
no_endpoint_a <- readRDS("../look_ups/no_endpoint_a.rds")
no_endpoint_b <- readRDS("../look_ups/no_endpoint_b.rds")

### eligible population ---------
# This includes individual respondents aged 16-64; Understanding Society waves 3-10 
eligible_a <- readRDS("../working_data/eligible_pop_a.rds") 
eligible_b <- readRDS("../working_data/eligible_pop_b.rds") 

### censored cases ---------
censor_combined_a <- readRDS("../look_ups/censor_combined_a.rds")
censor_combined_b <- readRDS("../look_ups/censor_combined_b.rds")

### analytic samples  ---------
dfas1a <- read_rds("../analytic_sample_data/dfas1a.rds")
dfas1b <- read_rds("../analytic_sample_data/dfas1b.rds")


##### flowchart nodes A to xx --------------------------------------------------

### Node A:  ----------------
### Total number of individuals and observations in study period
## number of unique individuals (node A)
raw_unique_indiv_a <- length(unique(master_raw_a$pidp))
raw_unique_indiv_b <- length(unique(master_raw_b$pidp))

## number of observations (node A)
raw_observations_a <- nrow(master_raw_a)
raw_observations_b <- nrow(master_raw_b)
#raw_observations <- prettyNum(raw_observations, big.mark=",")

### Node AA: ----------------
### number of individuals excluded due to not being working-age or no valid 
### response at study endpoint

## non-working age
non_work_age_a <- nrow(non_working_age_spine_a)
non_work_age_b <- nrow(non_working_age_spine_b)
#non_work_age <- prettyNum(non_work_age, big.mark=",")

## no valid endpoint response
no_valid_end_a <- length(unique(no_endpoint_a$pidp))
no_valid_end_b <- length(unique(no_endpoint_b$pidp))
#no_wv2 <- prettyNum(no_wv2, big.mark=",")


### Node B: ----------------
### number of individuals and observations in eligible population
eligible_unique_indiv_a <- length(unique(eligible_a$pidp))
eligible_unique_indiv_b <- length(unique(eligible_b$pidp))
#eligible_descr_unique_indiv <- prettyNum(eligible_descr_unique_indiv, big.mark=",")


### Node BB: ----------------
### number of individuals censored due to retirement or incomplete

## retired
retired_spine_a <- censor_combined_a %>% 
  filter(censor_reason=="retired")

retired_spine_b <- censor_combined_b %>% 
  filter(censor_reason=="retired")

censor_retired_a <- nrow(retired_spine_a)
censor_retired_b <- nrow(retired_spine_b)
#censor_retired <- prettyNum(censor_retired, big.mark=",")


## number of individuals censored due to incomplete data
incomplete_spine_a <- censor_combined_a %>% 
  filter(censor_reason=="incomplete")

incomplete_spine_b <- censor_combined_b %>% 
  filter(censor_reason=="incomplete")

incomplete_a <- nrow(incomplete_spine_a)
incomplete_b <- nrow(incomplete_spine_b)
#censor_lost <- prettyNum(censor_lost, big.mark=",")


### Node C: ---------------- 
### Number of individuals and observations in analytic samples

dfas1a_observations <- nrow(dfas1a)
dfas1b_observations <- nrow(dfas1b)
#dfas1_observations <- prettyNum(dfas1_observations, big.mark=",")

dfas1a_unique_indivs <- length(unique(dfas1a$pidp))
dfas1b_unique_indivs <- length(unique(dfas1b$pidp))
#dfas1_unique_indivs <- prettyNum(dfas1_unique_indivs, big.mark=",")


##### Study endpoint sample characteristics
sample_chars_endpoint <- readRDS("../working_data/sample_chars_endpoint.rds")  %>% 
  mutate(measure=tolower(measure))

## formatting for table
# for estimates set 1dp
sample_chars_endpoint$est <- formatC(sample_chars_endpoint$est, digits = 1, format = "f")

# for number of cases set 1,000 separator
sample_chars_endpoint$n <- prettyNum(sample_chars_endpoint$n, big.mark=",")

# for number of cases set <10 to -999 so can suppress low numbers
sample_chars_endpoint <- sample_chars_endpoint %>% 
  mutate(n = ifelse(n%in%c(1:10),-999,n))




sample_chars_endpoint_wide <- sample_chars_endpoint %>% 
  filter(wv_n==6) %>% 
  rename("n6" = "n",
         "est6" = "est") %>% 
  ungroup() %>% 
  select(-wv_n)

temp <- sample_chars_endpoint %>% 
  filter(wv_n==10) %>% 
  rename("n10" = "n",
         "est10" = "est") %>% 
  ungroup() %>% 
  select(-wv_n)

sample_chars_endpoint_wide <- sample_chars_endpoint_wide %>% 
  full_join(temp, by = c("var", "measure"))

# for number of cases convert to character so NAs can be removed
sample_chars_endpoint_wide$n6 <- as.character(sample_chars_endpoint_wide$n6)
sample_chars_endpoint_wide$n10 <- as.character(sample_chars_endpoint_wide$n10)

sample_chars_endpoint_wide <- sample_chars_endpoint_wide %>% 
  group_by(var) %>% 
  mutate(row_n = row_number()) %>% 
  mutate(var = ifelse(row_n==1,var,"")) %>% 
  ungroup() %>% 
  mutate(n6 = ifelse(n6=="NA","",
                    ifelse(n6=="-999","-",n6)),
         n10 = ifelse(n10=="NA","",
                    ifelse(n10=="-999","-",n10))) %>% 
  select(-row_n)

#### Node E: -------------
### Outcome caseness

srh_case <- "x"
srh_nocase <- "x" 
ghq_case <- "x"
ghq_nocase <- "x"

```

## Sample description

A total of `r prettyNum(raw_unique_indiv_a, big.mark=",")` individuals participated in UKHLS waves 3-6 and `r prettyNum(raw_unique_indiv_a, big.mark=",")` in waves 7-10 (Figure 1). We excluded `r prettyNum(non_work_age_a, big.mark=",")` and `r prettyNum(non_work_age_b, big.mark=",")` individuals respectively for not being aged 20-64 at the end point of the study period; plus `r prettyNum(no_valid_end_a, big.mark=",")` and `r prettyNum(no_valid_end_b, big.mark=",")` individuals for not having a valid survey response at wave 6 or 10 respectively. From eligible populations of `r prettyNum(eligible_unique_indiv_a, big.mark=",")` and `r prettyNum(eligible_unique_indiv_b, big.mark=",")` we excluded `r prettyNum(censor_retired_a, big.mark=",")` and `r prettyNum(censor_retired_a, big.mark=",")` retired individuals; and `r prettyNum(incomplete_a, big.mark=",")` and  `r prettyNum(incomplete_b, big.mark=",")` individuals who reported missing data for key exposure and outcomes of interest at waves 6 or 10 respectively. Table 1 presents key characteristics of the analytic sample at baseline by employment contract status (UKHLS wave 2).


**Figure 1: Flowchart for inclusion and exclusion of UKHLS respondents (waves 2-10) to create analytic sample**
```{r, echo=FALSE, include=TRUE, fig.width=8}


## flowchart A
grViz(paste0("digraph {
  graph []
  node []
    A [label = '",prettyNum(raw_unique_indiv_a, big.mark=",")," individuals
    
    (",prettyNum(raw_observations_a, big.mark=",")," observations)', shape = box, width = 3, height = 1]
    AA [label = '",prettyNum(non_work_age_a, big.mark=",")," non-working age
    
    ",prettyNum(no_valid_end_a, big.mark=",")," no valid wave 6 response' , shape = box, 
    width = 3, height = 1]
    B [label = '",prettyNum(eligible_unique_indiv_a, big.mark=",")," eligible individuals' , shape = box, 
    width = 3, height = 1]
    BB [label = '",prettyNum(censor_retired_a, big.mark=",")," retired
    
    ",prettyNum(incomplete_a, big.mark=",")," incomplete data', shape = box, width = 3, height = 1]
    C [label = '",prettyNum(dfas1a_unique_indivs, big.mark=",")," in analytic sample A
    
    (",prettyNum(dfas1a_observations, big.mark=",")," observations)', shape = box, 
    width = 3, height = 1]
    D [label = 'x in each employment history cluster' , shape = box, 
    width = 3, height = 1]
    E [label = '",prettyNum(srh_case, big.mark=",")," cases - poor self-reported health
    
    ",prettyNum(srh_nocase, big.mark=",")," event free - poor self-reported health
    
    ",prettyNum(ghq_case, big.mark=",")," cases - common mental health disorder
    
    ",prettyNum(ghq_nocase, big.mark=",")," event free - common mental health disorder' , shape = box, width = 3, height = 1]
  edge []
    A->B
    A->AA
    B->BB
    B->C
    C->D
    D->E
{ rank = same; A; AA }
{ rank = same; B; BB }}"))
```
```{r, echo=FALSE, include=TRUE, fig.width=8}


## flowchart B
grViz(paste0("digraph {
  graph []
  node []
    A [label = '",prettyNum(raw_unique_indiv_b, big.mark=",")," individuals
    
    (",prettyNum(raw_observations_b, big.mark=",")," observations)', shape = box, width = 3, height = 1]
    AA [label = '",prettyNum(non_work_age_b, big.mark=",")," non-working age
    
    ",prettyNum(no_valid_end_b, big.mark=",")," no valid wave 10 response' , shape = box, 
    width = 3, height = 1]
    B [label = '",prettyNum(eligible_unique_indiv_b, big.mark=",")," eligible individuals' , shape = box, 
    width = 3, height = 1]
    BB [label = '",prettyNum(censor_retired_b, big.mark=",")," retired
    
    ",prettyNum(incomplete_b, big.mark=",")," incomplete data', shape = box, width = 3, height = 1]
    C [label = '",prettyNum(dfas1b_unique_indivs, big.mark=",")," in analytic sample A
    
    (",prettyNum(dfas1b_observations, big.mark=",")," observations)', shape = box, 
    width = 3, height = 1]
    D [label = 'x in each employment history cluster' , shape = box, 
    width = 3, height = 1]
    E [label = '",prettyNum(srh_case, big.mark=",")," cases - poor self-reported health
    
    ",prettyNum(srh_nocase, big.mark=",")," event free - poor self-reported health
    
    ",prettyNum(ghq_case, big.mark=",")," cases - common mental health disorder
    
    ",prettyNum(ghq_nocase, big.mark=",")," event free - common mental health disorder' , shape = box, width = 3, height = 1]
  edge []
    A->B
    A->AA
    B->BB
    B->C
    C->D
    D->E
{ rank = same; A; AA }
{ rank = same; B; BB }}"))
```

```{r, echo=FALSE}
## add checks to ensure number of individuals adds up
check_AB_a <- eligible_unique_indiv_a+non_work_age_a+no_valid_end_a==raw_unique_indiv_a
check_BC_a <- censor_retired_a+incomplete_a+dfas1a_unique_indivs==eligible_unique_indiv_a

check_AB_b <- eligible_unique_indiv_b+non_work_age_b+no_valid_end_b==raw_unique_indiv_b
check_BC_b <- censor_retired_b+incomplete_b+dfas1b_unique_indivs==eligible_unique_indiv_b

if(check_AB_a!=TRUE){print("Sample A - boxes AA and B to not equal box A")}
if(check_BC_a!=TRUE){print("Sample A - boxes BB and C to not equal box B")}

if(check_AB_b!=TRUE){print("Sample B - boxes AA and B to not equal box A")}
if(check_BC_b!=TRUE){print("Sample B - boxes BB and C to not equal box B")}

```
\newpage
**Table 1: Analytic sample characteristics at baseline (UKHLS waves 6 and 10)**
```{r, echo=FALSE}

#sample_chars_endpoint_wide <- sample_chars_endpoint_wide[,c(1:10)]
kable(sample_chars_endpoint_wide, 
      col.names = c("Variable", 
                    "Measure", 
                    "n (wave 6)", 
                    "% (wave 6)", 
                    "n (wave 10)", 
                    "% (wave 10)"), 
      format.args = list(big.mark = ","))

```
\newpage
## Missing data
```{r, include=FALSE, echo=FALSE}
### total number of cases in each wave 16-64
## waves 3-6
spine_total_a <- dfas1a %>% group_by(wv_n) %>% 
  summarise(n=n()) %>% 
  ungroup() 

# number and % of individuals in sample based on endpoint
indivs_a <- spine_total_a[[4,2]]

spine_total_a <- spine_total_a %>% 
  mutate(total = indivs_a) %>% 
  mutate(pc = n/total*100) %>% 
  mutate(as = "a")

## waves7-10
spine_total_b <- dfas1b %>% group_by(wv_n) %>% 
  summarise(n=n()) %>% 
  ungroup() 

# number and % of individuals in sample based on endpoint
indivs_b <- spine_total_b[[4,2]]

spine_total_b <- spine_total_b %>% 
  mutate(total = indivs_b) %>% 
  mutate(pc = n/total*100) %>% 
  mutate(as = "b")

spine_total <- spine_total_a %>% bind_rows(spine_total_b)

### sequence analysis - wave non-response

## specify labels and codes (same for both samples)
spine_labs <- c("response", "missing")
spine_code <- c("RESP", "NA")


## wide format for sequencing data
# waves 3-6
spine_wide_a <- dfas1a %>%
  select(pidp,wv_n) %>% 
  mutate(response=1) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = response, values_fill = 99)

# waves 7-10
spine_wide_b <- dfas1b %>%
  select(pidp,wv_n) %>% 
  mutate(response=1) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = response, values_fill = 99)


spine.seq.a <- seqdef(spine_wide_a, 2:5, states = spine_code,
                    labels = spine_labs)
spine.seq.b <- seqdef(spine_wide_b, 2:5, states = spine_code,
                    labels = spine_labs)


### total number of cases by number of waves reported
## waves 3-6
spine_waves_a <- dfas1a %>% 
  group_by(pidp) %>% 
  summarise(n_wvs = n()) %>% # calculate number of waves responded to
  ungroup() %>% 
  group_by(n_wvs) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(pc= n/sum(n)*100) %>% 
  mutate(as = "a")

## waves 7-10
spine_waves_b <- dfas1b %>% 
  group_by(pidp) %>% 
  summarise(n_wvs = n()) %>% # calculate number of waves responded to
  ungroup() %>% 
  group_by(n_wvs) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(pc= n/sum(n)*100) %>% 
  mutate(as = "b")

```
Responses to all four waves in the study period were provided by `r spine_waves_a$pc[spine_waves_a$n_wvs==9]`% of analytic sample A and `r spine_waves_a$pc[spine_waves_a$n_wvs==9]`% of analytic sample B (Figure 2). Response sequences ascending in order of frequency are presented in Figure 3.

**Figure 2: Number of waves responded to by percentage of the analytic sample**
```{r, echo=FALSE, fig.width=3,fig.height=3}
## join together
spine_waves <- spine_waves_a %>% bind_rows(spine_waves_b)

# chart
spine_waves %>% 
  ggplot(aes(x=n_wvs, y=pc)) +
  geom_col() +
#  scale_x_continuous(limits = c(0,5),breaks = 1:4) +
  xlab("Number of waves responded to") +
  ylab("Percentage of cases") +
  theme_classic() +
  facet_wrap(~as, scales = "free")


```

\newpage
**Figure 3: UKHLS response sequences ascending in order of frequency (waves 3-10)**

```{r, echo=FALSE, fig.width=12,fig.height=15}

# sequence frequency plot (all common sequences)
#png("output/descriptive/seqfplot_missing.png", width = 960, height = 1920)
par(mfrow=c(1,3))
seqfplot(spine.seq.a, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "a")
seqfplot(spine.seq.b, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "b")
seqlegend(spine.seq.a, cex = 1.3)
#dev.off()
```

```{r, echo=FALSE, include=FALSE}

### recode employment status variables to create an employment contract variable
dfas1a_seq <- dfas1a %>% 
  mutate(emp_contract = ifelse(jbterm1 %in% c("not permanent job", 
                                              "or is there some way that it is not permanent?",
                                              "Or is there some way that it is not permanent?"),
                               "fixed-term",
                               ifelse(jbterm1 %in% c("a permanent job",
                                                     "A permanent job"),
                                      "permanent",
                                      ifelse(jbstat %in% c("Unemployed",
                                                           "unemployed"),
                                             "unemployed",
                                             ifelse(jbstat %in% c("on maternity leave",
                                                                  "On maternity leave",
                                                                  "Family care or home",
                                                    "full-time student",
                                                    "Full-time student",
                                                    "LT sick or disabled",
                                                    "Govt training scheme", # or fixed-term?
                                                    "On apprenticeship",    # or fixed-term?
                                                    "Unpaid, family business",
                                                    "doing something else",
                                                    "Doing something else"), 
                                                    "not in employment",
                                                    "missing")))))

dfas1b_seq <- dfas1b %>% 
  mutate(emp_contract = ifelse(jbterm1 %in% c("not permanent job", 
                                              "or is there some way that it is not permanent?",
                                              "Or is there some way that it is not permanent?"),
                               "fixed-term",
                               ifelse(jbterm1 %in% c("a permanent job",
                                                     "A permanent job"),
                                      "permanent",
                                      ifelse(jbstat %in% c("Unemployed",
                                                           "unemployed"),
                                             "unemployed",
                                             ifelse(jbstat %in% c("on maternity leave",
                                                                  "On maternity leave",
                                                                  "Family care or home",
                                                                  "full-time student",
                                                                  "Full-time student",
                                                                  "LT sick or disabled",
                                                                  "Govt training scheme", # or fixed-term?
                                                                  "On apprenticeship",    # or fixed-term?
                                                                  "Unpaid, family business",
                                                                  "doing something else",
                                                                  "Doing something else"), 
                                                    "not in employment",
                                                    "missing")))))

### wide format for creating sequence data
dfas1a_seq_wide  <-  dfas1a_seq %>% 
  select(pidp,wv_n,emp_contract) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = emp_contract, values_fill = "missing")

dfas1b_seq_wide  <-  dfas1b_seq %>% 
  select(pidp,wv_n,emp_contract) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = emp_contract, values_fill = "missing")

### define labels and codes for sequence analysis
## retaining missing values for now but plan to imput
emp_contract_labs <- c("fixed term","missing", "not in employment", 
                "permanent", "unemployed" )
emp_contract_code <- c("FT", "NE", "PE", "UE", "NA")

### create sequence data
emp_contract.seq.a <- seqdef(dfas1a_seq_wide, 2:5, states = emp_contract_code,
                    labels = emp_contract_labs)

emp_contract.seq.b <- seqdef(dfas1b_seq_wide, 2:5, states = emp_contract_code,
                             labels = emp_contract_labs)
```


```{r, echo=FALSE, fig.width=9,fig.height=9}

#
# sequence frequency plot (all common sequences)
par(mfrow=c(1,3))
seqfplot(emp_contract.seq.a, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sequence frequency plot")
seqfplot(emp_contract.seq.b, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sequence frequency plot")
seqlegend(emp_contract.seq.a, cex = 1.3)


```


```{r, echo=FALSE, include=FALSE}
#### ---------------------------------------------------------------------------
#### Broken employment
#### ---------------------------------------------------------------------------

### employment spells -----
# create numeric version
dfas1a_seq2 <- dfas1a %>% 
  mutate(nmpsp_dv=ifelse(nmpsp_dv=="none","0",nmpsp_dv)) %>% 
  mutate(nmpsp_dvn=as.numeric(nmpsp_dv))

dfas1b_seq2 <- dfas1b %>% 
  mutate(nmpsp_dv=ifelse(nmpsp_dv=="none","0",nmpsp_dv)) %>% 
  mutate(nmpsp_dvn=as.numeric(nmpsp_dv))

# create binary version (one spell or more)
dfas1a_seq2 <- dfas1a_seq2 %>% 
  mutate(emp_spells_bin=ifelse(nmpsp_dvn<1,"no",
                        ifelse(nmpsp_dvn>=1,"yes", NA)))

dfas1b_seq2 <- dfas1b_seq2 %>% 
  mutate(emp_spells_bin=ifelse(nmpsp_dvn<1,"no",
                               ifelse(nmpsp_dvn>=1,"yes", NA)))

### non-employment spells -----
# create numeric version
dfas1a_seq2 <- dfas1a_seq2 %>% 
  mutate(nnmpsp_dv=ifelse(nnmpsp_dv=="none","0",nnmpsp_dv)) %>% 
  mutate(nnmpsp_dvn=as.numeric(nnmpsp_dv))

dfas1b_seq2 <- dfas1b_seq2 %>% 
  mutate(nnmpsp_dv=ifelse(nnmpsp_dv=="none","0",nnmpsp_dv)) %>% 
  mutate(nnmpsp_dvn=as.numeric(nnmpsp_dv))

# create binary version (one or more spell)
dfas1a_seq2 <- dfas1a_seq2 %>% mutate(nonemp_spells_bin=ifelse(nnmpsp_dvn<1,"no",
                                                           ifelse(nnmpsp_dvn>=1,"yes", 
                                                                  NA)))

dfas1b_seq2 <- dfas1b_seq2 %>% mutate(nonemp_spells_bin=ifelse(nnmpsp_dvn<1,"no",
                                                               ifelse(nnmpsp_dvn>=1,"yes", 
                                                                      NA)))

### unemployment spells -----
# create numeric version
dfas1a_seq2 <- dfas1a_seq2 %>% 
  mutate(nunmpsp_dv=ifelse(nunmpsp_dv=="none","0",nunmpsp_dv)) %>% 
  mutate(nunmpsp_dvn=as.numeric(nunmpsp_dv))

dfas1b_seq2 <- dfas1b_seq2 %>% 
  mutate(nunmpsp_dv=ifelse(nunmpsp_dv=="none","0",nunmpsp_dv)) %>% 
  mutate(nunmpsp_dvn=as.numeric(nunmpsp_dv))

# create binary version (one or more spell)
dfas1a_seq2 <- dfas1a_seq2 %>% mutate(unemp_spells_bin=ifelse(nunmpsp_dvn<1,"no",
                                                          ifelse(nunmpsp_dvn>=1,"yes", 
                                                                 NA)))

dfas1b_seq2 <- dfas1b_seq2 %>% mutate(unemp_spells_bin=ifelse(nunmpsp_dvn<1,"no",
                                                              ifelse(nunmpsp_dvn>=1,"yes", 
                                                                     NA)))

### create broken employment variable ------
dfas1a_seq2 <- dfas1a_seq2 %>% 
  mutate(broken_emp = ifelse(emp_spells_bin=="no","No employment spells", 
                             ifelse(emp_spells_bin=="yes" & 
                                      nonemp_spells_bin=="no" & 
                                      unemp_spells_bin=="no",
                                    "Unbroken employment",
                                    ifelse(emp_spells_bin=="yes" & 
                                             (nonemp_spells_bin=="yes" |
                                                unemp_spells_bin=="yes"),
                                           "Broken employment","check"))))

dfas1b_seq2 <- dfas1b_seq2 %>% 
  mutate(broken_emp = ifelse(emp_spells_bin=="no","No employment spells", 
                             ifelse(emp_spells_bin=="yes" & 
                                      nonemp_spells_bin=="no" & 
                                      unemp_spells_bin=="no",
                                    "Unbroken employment",
                                    ifelse(emp_spells_bin=="yes" & 
                                             (nonemp_spells_bin=="yes" |
                                                unemp_spells_bin=="yes"),
                                           "Broken employment","check"))))

### wide format for creating sequence data
dfas1a_seq_wide2  <-  dfas1a_seq2 %>% 
  select(pidp,wv_n,broken_emp) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = broken_emp, values_fill = "missing")

dfas1b_seq_wide2  <-  dfas1b_seq2 %>% 
  select(pidp,wv_n,broken_emp) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = broken_emp, values_fill = "missing")

### define labels and codes for sequence analysis
## retaining missing values for now but plan to impute
broken_emp_labs <- c("Broken employment", "Missing", "No employment spells", "Unbroken employment" )
broken_emp_code <- c("BE","NA", "NE", "UE")

### create sequence data
broken_emp.seq.a <- seqdef(dfas1a_seq_wide2, 2:5, states = broken_emp_code,
                             labels = broken_emp_labs)

broken_emp.seq.b <- seqdef(dfas1b_seq_wide2, 2:5, states = broken_emp_code,
                             labels = broken_emp_labs)

```


```{r, echo=FALSE, fig.width=9,fig.height=9}
# sequence frequency plot (all common sequences)
#png("output/descriptive/seqfplot_emp_contract.png", width = 960, height = 1920)
par(mfrow=c(1,3))
seqfplot(broken_emp.seq.a, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sequence frequency plot")
seqfplot(broken_emp.seq.b, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sequence frequency plot")
seqlegend(broken_emp.seq.a, cex = 1.3)

```

```{r, echo=FALSE, include=FALSE}

#### ---------------------------------------------------------------------------
#### Multiple jobs
#### ---------------------------------------------------------------------------

## collapse missing categories
dfas1a$j2has <- as.character(dfas1a$j2has)
dfas1b$j2has <- as.character(dfas1b$j2has)


dfas1a_seq3 <- dfas1a %>% 
  mutate(j2has_dv = ifelse(j2has %in% c("proxy","missing","refusal","don't know"),
                           "missing", 
                           ifelse(jbstat %in% c("Unemployed", "unemployed"),
                                             "unemployed",
                          ifelse(jbstat %in% c("on maternity leave",
                                               "On maternity leave",
                                               "Family care or home",
                                               "full-time student",
                                               "Full-time student",
                                               "LT sick or disabled",
                                               "Govt training scheme", # or fixed-term?
                                               "On apprenticeship",    # or fixed-term?
                                               "Unpaid, family business",
                                               "doing something else",
                                               "Doing something else"), 
                                                    "not in employment",
                                                    j2has))))

dfas1b_seq3 <- dfas1b %>% 
  mutate(j2has_dv = ifelse(j2has %in% c("proxy","missing","refusal","don't know"),
                           "missing", 
                           ifelse(jbstat %in% c("Unemployed", "unemployed"),
                                  "unemployed",
                                  ifelse(jbstat %in% c("on maternity leave",
                                                       "On maternity leave",
                                                       "Family care or home",
                                                       "full-time student",
                                                       "Full-time student",
                                                       "LT sick or disabled",
                                                       "Govt training scheme", # or fixed-term?
                                                       "On apprenticeship",    # or fixed-term?
                                                       "Unpaid, family business",
                                                       "doing something else",
                                                       "Doing something else"), 
                                         "not in employment",
                                         j2has))))


### wide format for creating sequence data
dfas1a_seq_wide3  <-  dfas1a_seq3 %>% 
  select(pidp,wv_n,j2has_dv) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = j2has_dv, values_fill = "missing")

dfas1b_seq_wide3  <-  dfas1b_seq3 %>% 
  select(pidp,wv_n,j2has_dv) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = j2has_dv, values_fill = "missing")

### define labels and codes for sequence analysis
## retaining missing values for now but plan to impute
multi_jobs_labs <- c("Missing", "No", "not in employment", "unemployed", "Yes")
multi_jobs_code <- c("NA","N", "NE", "UE", "Y")

### create sequence data
multi_jobs.seq.a <- seqdef(dfas1a_seq_wide3, 2:5, states = multi_jobs_code,
                           labels = multi_jobs_labs)

multi_jobs.seq.b <- seqdef(dfas1b_seq_wide3, 2:5, states = multi_jobs_code,
                           labels = multi_jobs_labs)
```

```{r, echo=FALSE, fig.width=9,fig.height=9}
# sequence frequency plot (all common sequences)
#png("output/descriptive/seqfplot_emp_contract.png", width = 960, height = 1920)
par(mfrow=c(1,3))
seqfplot(multi_jobs.seq.a, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sequence frequency plot")
seqfplot(multi_jobs.seq.b, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sequence frequency plot")
seqlegend(multi_jobs.seq.a, cex = 1.3)
```