---
title: "Precarious employment - descriptive analysis"
author: "Andrew Pulford"
date: ""
output: word_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## remove any existing objects from global environment
rm(list=ls()) 

## stop scientific notation
options(scipen=999, digits = 1)

### load packages
library(tidyverse)
library(DiagrammeR)
library(knitr)
library(TraMineR)
remotes::install_github("rstudio/webshot2")

##### load dataframes ----------------------------------------------------------

### master raw data --------
# This includes selected variables for all individual respondents across Understanding Society waves 3-10
master_raw <- readRDS("../raw_data/master_raw1.rds")

## waves 3-6
master_raw_a <- master_raw %>% filter(wv_n %in% c(3:6))

## waves 7-10
master_raw_b <- master_raw %>% filter(wv_n %in% c(7:10))

### non-working age ---------
non_working_age_spine_a <- readRDS("../look_ups/non_working_age_spine_a.rds")
non_working_age_spine_b <- readRDS("../look_ups/non_working_age_spine_b.rds")

### no valid case at end-point ----------
no_endpoint_a <- readRDS("../look_ups/no_endpoint_a.rds")
no_endpoint_b <- readRDS("../look_ups/no_endpoint_b.rds")

### eligible population ---------
# This includes individual respondents aged 16-64; Understanding Society waves 3-10 
eligible_a <- readRDS("../working_data/eligible_pop_a.rds") 
eligible_b <- readRDS("../working_data/eligible_pop_b.rds") 

### censored cases ---------
censor_combined_a <- readRDS("../look_ups/censor_combined_a.rds")
censor_combined_b <- readRDS("../look_ups/censor_combined_b.rds")

### analytic samples  ---------
dfas1a <- read_rds("../analytic_sample_data/dfas1a.rds")
dfas1b <- read_rds("../analytic_sample_data/dfas1b.rds")

### a priori exposure groupings
emp_contract_group_df <- readRDS("../working_data/emp_contract_group_df.rds")
emp_spells_group_df <- readRDS("../working_data/emp_spells_group_df.rds")
multi_jobs_group_df<- readRDS("../working_data/multi_jobs_group_df.rds")

##### flowchart nodes A to xx --------------------------------------------------

### Node A:  ----------------
### Total number of individuals and observations in study period
## number of unique individuals (node A)
raw_unique_indiv_a <- length(unique(master_raw_a$pidp))
raw_unique_indiv_b <- length(unique(master_raw_b$pidp))

## number of observations (node A)
raw_observations_a <- nrow(master_raw_a)
raw_observations_b <- nrow(master_raw_b)
#raw_observations <- prettyNum(raw_observations, big.mark=",")

### Node AA: ----------------
### number of individuals excluded due to not being working-age or no valid 
### response at study endpoint

## non-working age
non_work_age_a <- nrow(non_working_age_spine_a)
non_work_age_b <- nrow(non_working_age_spine_b)
#non_work_age <- prettyNum(non_work_age, big.mark=",")

## no valid endpoint response
no_valid_end_a <- length(unique(no_endpoint_a$pidp))
no_valid_end_b <- length(unique(no_endpoint_b$pidp))
#no_wv2 <- prettyNum(no_wv2, big.mark=",")


### Node B: ----------------
### number of individuals and observations in eligible population
eligible_unique_indiv_a <- length(unique(eligible_a$pidp))
eligible_unique_indiv_b <- length(unique(eligible_b$pidp))
#eligible_descr_unique_indiv <- prettyNum(eligible_descr_unique_indiv, big.mark=",")


### Node BB: ----------------
### number of individuals censored due to retirement or incomplete

## retired
retired_spine_a <- censor_combined_a %>% 
  filter(censor_reason=="retired")

retired_spine_b <- censor_combined_b %>% 
  filter(censor_reason=="retired")

censor_retired_a <- nrow(retired_spine_a)
censor_retired_b <- nrow(retired_spine_b)
#censor_retired <- prettyNum(censor_retired, big.mark=",")


## number of individuals censored due to incomplete data
incomplete_spine_a <- censor_combined_a %>% 
  filter(censor_reason=="incomplete")

incomplete_spine_b <- censor_combined_b %>% 
  filter(censor_reason=="incomplete")

incomplete_a <- nrow(incomplete_spine_a)
incomplete_b <- nrow(incomplete_spine_b)
#censor_lost <- prettyNum(censor_lost, big.mark=",")


### Node C: ---------------- 
### Number of individuals and observations in analytic samples

dfas1a_observations <- nrow(dfas1a)
dfas1b_observations <- nrow(dfas1b)
#dfas1_observations <- prettyNum(dfas1_observations, big.mark=",")

dfas1a_unique_indivs <- length(unique(dfas1a$pidp))
dfas1b_unique_indivs <- length(unique(dfas1b$pidp))
#dfas1_unique_indivs <- prettyNum(dfas1_unique_indivs, big.mark=",")


##### Study endpoint sample characteristics
sample_chars_endpoint <- readRDS("../working_data/sample_chars_endpoint.rds")  %>% 
  mutate(measure=tolower(measure))

## formatting for table
# for estimates set 1dp
sample_chars_endpoint$est <- formatC(sample_chars_endpoint$est, digits = 1, format = "f")

# for number of cases set 1,000 separator
sample_chars_endpoint$n <- prettyNum(sample_chars_endpoint$n, big.mark=",")

# for number of cases set <10 to -999 so can suppress low numbers
sample_chars_endpoint <- sample_chars_endpoint %>% 
  mutate(n = ifelse(n%in%c(1:10),-999,n))




sample_chars_endpoint_wide <- sample_chars_endpoint %>% 
  filter(wv_n==6) %>% 
  rename("n6" = "n",
         "est6" = "est") %>% 
  ungroup() %>% 
  select(-wv_n)

temp <- sample_chars_endpoint %>% 
  filter(wv_n==10) %>% 
  rename("n10" = "n",
         "est10" = "est") %>% 
  ungroup() %>% 
  select(-wv_n)

sample_chars_endpoint_wide <- sample_chars_endpoint_wide %>% 
  full_join(temp, by = c("var", "measure"))

# for number of cases convert to character so NAs can be removed
sample_chars_endpoint_wide$n6 <- as.character(sample_chars_endpoint_wide$n6)
sample_chars_endpoint_wide$n10 <- as.character(sample_chars_endpoint_wide$n10)

sample_chars_endpoint_wide <- sample_chars_endpoint_wide %>% 
  group_by(var) %>% 
  mutate(row_n = row_number()) %>% 
  mutate(var = ifelse(row_n==1,var,"")) %>% 
  ungroup() %>% 
  mutate(n6 = ifelse(n6=="NA","",
                    ifelse(n6=="-999","-",n6)),
         n10 = ifelse(n10=="NA","",
                    ifelse(n10=="-999","-",n10))) %>% 
  select(-row_n)

#### Node D: -------------
### Exposure groupings

# to be added

#### Node E: -------------
### Outcome caseness

## self-rated health

# events
srh_case <- "x"
srh_case_a <- dfas1a %>% filter(wv_n==6 & srh_bin=="good/fair/poor")
srh_case_a <- srh_case_a %>% nrow()
  
srh_case_b <- dfas1b %>% filter(wv_n==10 & srh_bin=="good/fair/poor")
srh_case_b <- srh_case_b %>% nrow()

# non-events
srh_nocase <- "x"
srh_nocase_a <- dfas1a %>% filter(wv_n==6 & srh_bin=="excellent/very good")
srh_nocase_a <- srh_nocase_a %>% nrow()
  
srh_nocase_b <- dfas1b %>% filter(wv_n==10 & srh_bin=="excellent/very good")
srh_nocase_b <- srh_nocase_b %>% nrow()
## GHQ-12
# events
ghq_case <- "x"

ghq_case_a <- dfas1a %>% filter(wv_n==6 & ghq_case3=="3 or more")
ghq_case_a <- ghq_case_a %>% nrow()

ghq_case_b <- dfas1b %>% filter(wv_n==10 & ghq_case3=="3 or more")
ghq_case_b <- ghq_case_b %>% nrow()

# non-events
ghq_nocase <- "x"
ghq_nocase_a <- dfas1a %>% filter(wv_n==6 & ghq_case3=="0-2")
ghq_nocase_a <- ghq_nocase_a %>% nrow()

ghq_nocase_b <- dfas1b %>% filter(wv_n==10 & ghq_case3=="0-2")
ghq_nocase_b <- ghq_nocase_b %>% nrow()

```

*UNWEIGHTED RESULTS - FOR ILLUSTRATIVE PURPOSES ONLY*

##Sample description

A total of `r prettyNum(raw_unique_indiv_a, big.mark=",")` individuals participated in UKHLS waves 3-6 and `r prettyNum(raw_unique_indiv_a, big.mark=",")` in waves 7-10 (Figure 1). We excluded `r prettyNum(non_work_age_a, big.mark=",")` and `r prettyNum(non_work_age_b, big.mark=",")` individuals respectively for not being aged 20-64 at the end point of the study period; plus `r prettyNum(no_valid_end_a, big.mark=",")` and `r prettyNum(no_valid_end_b, big.mark=",")` individuals for not having a valid survey response at wave 6 or 10 respectively. From eligible populations of `r prettyNum(eligible_unique_indiv_a, big.mark=",")` and `r prettyNum(eligible_unique_indiv_b, big.mark=",")` we excluded `r prettyNum(censor_retired_a, big.mark=",")` and `r prettyNum(censor_retired_b, big.mark=",")` retired individuals; and `r prettyNum(incomplete_a, big.mark=",")` and  `r prettyNum(incomplete_b, big.mark=",")` individuals who reported missing data for key exposure and outcomes of interest at waves 6 or 10 respectively. Table 1 presents key characteristics of the analytic sample at baseline by employment contract status (UKHLS wave 2).


**Figure 1: Flowcharts for inclusion and exclusion of UKHLS respondents to create analytic sample**

**(a) Sample A (waves 3-6)**
```{r, echo=FALSE, include=TRUE, fig.width=8}


## flowchart A
grViz(paste0("digraph {
  graph []
  node []
    A [label = '",prettyNum(raw_unique_indiv_a, big.mark=",")," individuals
    
    (",prettyNum(raw_observations_a, big.mark=",")," observations)', shape = box, width = 3, height = 1]
    AA [label = '",prettyNum(non_work_age_a, big.mark=",")," non-working age
    
    ",prettyNum(no_valid_end_a, big.mark=",")," no valid wave 6 response' , shape = box, 
    width = 3, height = 1]
    B [label = '",prettyNum(eligible_unique_indiv_a, big.mark=",")," eligible individuals' , shape = box, 
    width = 3, height = 1]
    BB [label = '",prettyNum(censor_retired_a, big.mark=",")," retired
    
    ",prettyNum(incomplete_a, big.mark=",")," incomplete data', shape = box, width = 3, height = 1]
    C [label = '",prettyNum(dfas1a_unique_indivs, big.mark=",")," in analytic sample A
    
    (",prettyNum(dfas1a_observations, big.mark=",")," observations)', shape = box, 
    width = 3, height = 1]
    D [label = 'x in each employment history cluster' , shape = box, 
    width = 3, height = 1]
    E [label = '",prettyNum(srh_case_a, big.mark=",")," cases - poor self-reported health
    
    ",prettyNum(srh_nocase_a, big.mark=",")," event free - poor self-reported health
    
    ",prettyNum(ghq_case_a, big.mark=",")," cases - common mental health disorder
    
    ",prettyNum(ghq_nocase_a, big.mark=",")," event free - common mental health disorder' , shape = box, width = 3, height = 1]
  edge []
    A->B
    A->AA
    B->BB
    B->C
    C->D
    D->E
{ rank = same; A; AA }
{ rank = same; B; BB }}"))
```

**(b) Sample B (waves 7-10)**
```{r, echo=FALSE, include=TRUE, fig.width=8}


## flowchart B
grViz(paste0("digraph {
  graph []
  node []
    A [label = '",prettyNum(raw_unique_indiv_b, big.mark=",")," individuals
    
    (",prettyNum(raw_observations_b, big.mark=",")," observations)', shape = box, width = 3, height = 1]
    AA [label = '",prettyNum(non_work_age_b, big.mark=",")," non-working age
    
    ",prettyNum(no_valid_end_b, big.mark=",")," no valid wave 10 response' , shape = box, 
    width = 3, height = 1]
    B [label = '",prettyNum(eligible_unique_indiv_b, big.mark=",")," eligible individuals' , shape = box, 
    width = 3, height = 1]
    BB [label = '",prettyNum(censor_retired_b, big.mark=",")," retired
    
    ",prettyNum(incomplete_b, big.mark=",")," incomplete data', shape = box, width = 3, height = 1]
    C [label = '",prettyNum(dfas1b_unique_indivs, big.mark=",")," in analytic sample A
    
    (",prettyNum(dfas1b_observations, big.mark=",")," observations)', shape = box, 
    width = 3, height = 1]
    D [label = 'x in each employment history cluster' , shape = box, 
    width = 3, height = 1]
    E [label = '",prettyNum(srh_case_b, big.mark=",")," cases - poor self-reported health
    
    ",prettyNum(srh_nocase_b, big.mark=",")," event free - poor self-reported health
    
    ",prettyNum(ghq_case_b, big.mark=",")," cases - common mental health disorder
    
    ",prettyNum(ghq_nocase_b, big.mark=",")," event free - common mental health disorder' , shape = box, width = 3, height = 1]
  edge []
    A->B
    A->AA
    B->BB
    B->C
    C->D
    D->E
{ rank = same; A; AA }
{ rank = same; B; BB }}"))
```

```{r, echo=FALSE}
## add checks to ensure number of individuals adds up
check_AB_a <- eligible_unique_indiv_a+non_work_age_a+no_valid_end_a==raw_unique_indiv_a
check_BC_a <- censor_retired_a+incomplete_a+dfas1a_unique_indivs==eligible_unique_indiv_a

check_AB_b <- eligible_unique_indiv_b+non_work_age_b+no_valid_end_b==raw_unique_indiv_b
check_BC_b <- censor_retired_b+incomplete_b+dfas1b_unique_indivs==eligible_unique_indiv_b

if(check_AB_a!=TRUE){print("Sample A - boxes AA and B to not equal box A")}
if(check_BC_a!=TRUE){print("Sample A - boxes BB and C to not equal box B")}

if(check_AB_b!=TRUE){print("Sample B - boxes AA and B to not equal box A")}
if(check_BC_b!=TRUE){print("Sample B - boxes BB and C to not equal box B")}

```
\newpage
**Table 1: Analytic sample characteristics at study endpoint (UKHLS waves 6 and 10)**
```{r, echo=FALSE}

#sample_chars_endpoint_wide <- sample_chars_endpoint_wide[,c(1:10)]
kable(sample_chars_endpoint_wide, 
      col.names = c("Variable", 
                    "Measure", 
                    "n (wave 6)", 
                    "% (wave 6)", 
                    "n (wave 10)", 
                    "% (wave 10)"), 
      format.args = list(big.mark = ","))

```
\newpage
## Missing data
```{r, include=FALSE, echo=FALSE}
### total number of cases in each wave 16-64
## waves 3-6
spine_total_a <- dfas1a %>% group_by(wv_n) %>% 
  summarise(n=n()) %>% 
  ungroup() 

# number and % of individuals in sample based on endpoint
indivs_a <- spine_total_a[[4,2]]

spine_total_a <- spine_total_a %>% 
  mutate(total = indivs_a) %>% 
  mutate(pc = n/total*100) %>% 
  mutate(as = "a")

## waves7-10
spine_total_b <- dfas1b %>% group_by(wv_n) %>% 
  summarise(n=n()) %>% 
  ungroup() 

# number and % of individuals in sample based on endpoint
indivs_b <- spine_total_b[[4,2]]

spine_total_b <- spine_total_b %>% 
  mutate(total = indivs_b) %>% 
  mutate(pc = n/total*100) %>% 
  mutate(as = "b")

spine_total <- spine_total_a %>% bind_rows(spine_total_b)

### sequence analysis - wave non-response

## specify labels and codes (same for both samples)
spine_labs <- c("response", "missing")
spine_code <- c("RESP", "NA")


## wide format for sequencing data
# waves 3-6
spine_wide_a <- dfas1a %>%
  select(pidp,wv_n) %>% 
  mutate(response=1) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = response, values_fill = 99)

# waves 7-10
spine_wide_b <- dfas1b %>%
  select(pidp,wv_n) %>% 
  mutate(response=1) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = response, values_fill = 99)


spine.seq.a <- seqdef(spine_wide_a, 2:5, states = spine_code,
                    labels = spine_labs)
spine.seq.b <- seqdef(spine_wide_b, 2:5, states = spine_code,
                    labels = spine_labs)


### total number of cases by number of waves reported
## waves 3-6
spine_waves_a <- dfas1a %>% 
  group_by(pidp) %>% 
  summarise(n_wvs = n()) %>% # calculate number of waves responded to
  ungroup() %>% 
  group_by(n_wvs) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(pc= n/sum(n)*100) %>% 
  mutate(as = "Sample A (waves 3-6)")

## waves 7-10
spine_waves_b <- dfas1b %>% 
  group_by(pidp) %>% 
  summarise(n_wvs = n()) %>% # calculate number of waves responded to
  ungroup() %>% 
  group_by(n_wvs) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(pc= n/sum(n)*100) %>% 
  mutate(as = "Sample B (waves 7-10")

```
Responses to all four waves in the study period were provided by `r spine_waves_a$pc[spine_waves_a$n_wvs==9]`% of analytic sample A and `r spine_waves_a$pc[spine_waves_a$n_wvs==9]`% of analytic sample B (Figure 2). Response sequences ascending in order of frequency are presented in Figure 3.

**Figure 2: Number of waves responded to by percentage of the analytic sample**
```{r, echo=FALSE, fig.width=4,fig.height=3}
## join together
spine_waves <- spine_waves_a %>% bind_rows(spine_waves_b)

# chart
spine_waves %>% 
  ggplot(aes(x=n_wvs, y=pc)) +
  geom_col() +
#  scale_x_continuous(limits = c(0,5),breaks = 1:4) +
  xlab("Number of waves responded to") +
  ylab("Percentage of cases") +
  theme_classic() +
  facet_wrap(~as, scales = "free")


```

\newpage
**Figure 3: Valid response sequences ascending in order of frequency**

```{r, echo=FALSE, fig.width=12,fig.height=15}

# sequence frequency plot (all common sequences)
#png("output/descriptive/seqfplot_missing.png", width = 960, height = 1920)
par(mfrow=c(1,3))
seqfplot(spine.seq.a, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sample A (waves 3-6)")
seqfplot(spine.seq.b, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sample B (waves 7-10)")
seqlegend(spine.seq.a, cex = 1.3)
#dev.off()
```



```{r, echo=FALSE, include=FALSE}

### recode employment status variables to create an employment contract variable
dfas1a_seq <- dfas1a

dfas1b_seq <- dfas1b 

### wide format for creating sequence data
dfas1a_seq_wide  <-  dfas1a_seq %>% 
  select(pidp,wv_n,emp_contract) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = emp_contract, values_fill = "missing")

dfas1b_seq_wide  <-  dfas1b_seq %>% 
  select(pidp,wv_n,emp_contract) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = emp_contract, values_fill = "missing")

### define labels and codes for sequence analysis
## retaining missing values for now but plan to imput
emp_contract_labs <- c("fixed term", "missing", "permanent", "unemployed/not in employment" )
emp_contract_code <- c("FT", "NA", "PE", "UE")

### create sequence data
emp_contract.seq.a <- seqdef(dfas1a_seq_wide, 2:5, states = emp_contract_code,
                    labels = emp_contract_labs)

emp_contract.seq.b <- seqdef(dfas1b_seq_wide, 2:5, states = emp_contract_code,
                             labels = emp_contract_labs)

```

\newpage

## Sequence analysis of precrious employment exposures of interest

hhdgfd

### Employment contract

ndjfhdj

**Figure 4: Employment contract sequences ascending in order of frequency**
```{r, echo=FALSE, fig.width=9,fig.height=9}

#
# sequence frequency plot (all common sequences)
par(mfrow=c(1,3))
seqfplot(emp_contract.seq.a, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sample A (waves 3-6)")
seqfplot(emp_contract.seq.b, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sample B (waves 7-10)")
seqlegend(emp_contract.seq.a, cex = 1.3)


```


```{r, echo=FALSE, include=FALSE}
#### ---------------------------------------------------------------------------
#### Broken employment
#### ---------------------------------------------------------------------------

### employment spells -----
# create numeric version
dfas1a_seq2 <- dfas1a

dfas1b_seq2 <- dfas1b 

### wide format for creating sequence data
dfas1a_seq_wide2  <-  dfas1a_seq2 %>% 
  select(pidp,wv_n,broken_emp) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
#  mutate(broken_emp = ifelse(is.na(broken_emp),"missing",broken_emp)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = broken_emp, values_fill = "missing")

dfas1b_seq_wide2  <-  dfas1b_seq2 %>% 
  select(pidp,wv_n,broken_emp) %>% 
  mutate(broken_emp = ifelse(is.na(broken_emp),"missing",broken_emp)) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = broken_emp, values_fill = "missing")

### define labels and codes for sequence analysis
## retaining missing values for now but plan to impute
broken_emp_labs <- c("Broken employment", "Missing", "No employment spells", "Unbroken employment" )
broken_emp_code <- c("BE","NA", "NE", "UE")

### create sequence data
broken_emp.seq.a <- seqdef(dfas1a_seq_wide2, 2:5, states = broken_emp_code,
                             labels = broken_emp_labs)

broken_emp.seq.b <- seqdef(dfas1b_seq_wide2, 2:5, states = broken_emp_code,
                           labels = broken_emp_labs)

```

**Table x: **
```{r, echo=FALSE}
kable(emp_contract_group_df, 
      col.names = c("A priori employment contract grouping", 
                    "n (Sample A)", 
                    "% (Sample A)", 
                    "n (Sample B)", 
                    "% (Sample B)"), 
      format.args = list(big.mark = ","))

```


### Employment spells
hghgh


**Figure 5: Employment spell sequences ascending in order of frequency**
```{r, echo=FALSE, fig.width=9,fig.height=9}
# sequence frequency plot (all common sequences)
#png("output/descriptive/seqfplot_emp_contract.png", width = 960, height = 1920)
par(mfrow=c(1,3))
seqfplot(broken_emp.seq.a, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sample A (waves 3-6)")
seqfplot(broken_emp.seq.b, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sample A (waves 7-10)")
seqlegend(broken_emp.seq.a, cex = 1.3)

```

**Table x: **
```{r, echo=FALSE}
kable(emp_spells_group_df, 
      col.names = c("A priori broken employment spells grouping", 
                    "n (Sample A)", 
                    "% (Sample A)", 
                    "n (Sample B)", 
                    "% (Sample B)"), 
      format.args = list(big.mark = ","))

```

```{r, echo=FALSE, include=FALSE}

#### ---------------------------------------------------------------------------
#### Multiple jobs
#### ---------------------------------------------------------------------------

dfas1a_seq3 <- dfas1a
dfas1b_seq3 <- dfas1b

### wide format for creating sequence data
dfas1a_seq_wide3  <-  dfas1a_seq3 %>% 
  select(pidp,wv_n,j2has_dv) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = j2has_dv, values_fill = "missing")

dfas1b_seq_wide3  <-  dfas1b_seq3 %>% 
  select(pidp,wv_n,j2has_dv) %>% 
  mutate(wv=paste0("wv_",wv_n)) %>% 
  select(-wv_n) %>% 
  pivot_wider(names_from = wv, values_from = j2has_dv, values_fill = "missing")

### define labels and codes for sequence analysis
## retaining missing values for now but plan to impute
multi_jobs_labs <- c("Missing", "No", "Yes")
multi_jobs_code <- c("NA","N", "Y")

### create sequence data
multi_jobs.seq.a <- seqdef(dfas1a_seq_wide3, 2:5, states = multi_jobs_code,
                           labels = multi_jobs_labs)

multi_jobs.seq.b <- seqdef(dfas1b_seq_wide3, 2:5, states = multi_jobs_code,
                           labels = multi_jobs_labs)

```

### Multiple jobs


**Figure 6: Multiple job sequences ascending in order of frequency**
```{r, echo=FALSE, fig.width=9,fig.height=9}
# sequence frequency plot (all common sequences)
#png("output/descriptive/seqfplot_emp_contract.png", width = 960, height = 1920)
par(mfrow=c(1,3))
seqfplot(multi_jobs.seq.a, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sample A (waves 3-6)")
seqfplot(multi_jobs.seq.b, 
         idxs=1:900, # to add more lines
         with.legend = F, 
         border = NA, 
         main = "Sample B (waves 7-10)")
seqlegend(multi_jobs.seq.a, cex = 1.3)
```
**Table x: **
```{r, echo=FALSE}
kable(multi_jobs_group_df, 
      col.names = c("A priori multiple grouping", 
                    "n (Sample A)", 
                    "% (Sample A)", 
                    "n (Sample B)", 
                    "% (Sample B)"), 
      format.args = list(big.mark = ","))

```